Inefficient Concessions and Mediation

### 2017-09-09 draft.tex
remove differentiated costs (not necessary)

In period 0, countries 1 and 2 simultaneously give costly concessions $g_1$ and $g_2$ $\in \mathbb{R} \geq 0$. Concessions are given at cost $C(g_i)$ and benefit the recipient $g_i$, where $i\in\left\{l,h\right\}$. For simplicity, we initially assume $C(g_i)= g_i$. We will relax this assumption later.

@@

Welfare is again considered as the sum of the high type utilities. Qualitatively, similar results would hold if we also included low types in the analysis with weighting based on $p$. In the analysis, incentive compatibility and individual rationality constraints must be satisfied. The revelation principle is used to attain truthful self-identification (Myerson 1979). For mediation to be useful, $C_l(\cdot)$ needs to be pointwise greater than $C_h(\cdot)$, and we will assume this. This means that a low type is not identical to a high type, and indeed bears greater costs for peaceful actions. This can be thought of as a leader who must explain peace to his people. If the people want to make peace with the rival, posturing for peace is not costly for the leader. But if the people do not want to make peace with the rival, the leader loses credibility by outwardly touting peace. Therefore taking a peaceful action is more costly, and perhaps even impossible. This low type leader can bluff in the bilateral setting but not, as we will see, when a mediator is present. 
\\
\\
{\bf Theorem 5:} A mediator can be used to eliminate inefficient concessions. \\
\emph{Proof:}  $C_l(\cdot)> C_h(\cdot)$ as per the assumption above.  The low type IC ($pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h$) binds, so the equilibrium benchmark gift is $g^{*}=p(X_{FT}^l-X_{FF}^l)$ and high type utility is $U_h^*= pX_{TT}+(1-p)X_{FF} - (1-p)g^*$). Under $M$, the low type IC can bind without having high types give a concession to low types. The equilibrium separating gift $g^{*M}$ will be larger than $g^*$ in this case, specifically 
\be
g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF}
\ee
because high types exchange concessions of the same value. $U_h^M>U_h^*$. Note that no additional wealth is created by $M$, but less is transferred to low types, and welfare is defined as high-type utility.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$}  $\blacksquare$

Since the low type IC constraint binds, the only way in which a mechanism can improve upon the benchmark case is by having asymmetric costs of giving for the types. As per Theorem 4, inefficiency is due to the high types giving an inefficient gift in order to separate while still being able to prevent low types from using concessions against the high types. If costs are disparate, under certain parameters the mediator can mandate concessions only between declared high types. This means there is no longer a need for inefficient concessions since concessions will never be used against the giver. Thus, a mediator is able to remove inefficiency in concession giving and bring willing parties to the table to give the gifts necessary to achieve peace.

>>>>

In period 0, countries 1 and 2 simultaneously give costly concessions $g_1$ and $g_2$ $\in \mathbb{R} \geq 0$. Concessions are given at cost $C(g_i)$ and benefit the recipient in the amount $g_i$, where $i\in\left\{1,2\right\}$. For simplicity, we assume $C(g_i)= g_i$.

@@

Welfare is again considered as the sum of the high type utilities. Qualitatively, similar results would hold if we also included low types in the analysis with weighting based on $p$. In the analysis, incentive compatibility and individual rationality constraints must be satisfied. The revelation principle is used to attain truthful self-identification (Myerson 1979).

\\
{\bf Theorem 5:} A mediator can be used to eliminate inefficient concessions. \\
\emph{Proof:}  The low type IC ($pg_h +X_{FF}^l \geq pX_{FT}^l+(1-p)X_{FF}^l-g_h+pg_h$) binds, so the equilibrium benchmark gift is $g^{*}=p(X_{FT}^l-X_{FF}^l)$ and high type utility is $U_h^*= pX_{TT}+(1-p)X_{FF} - (1-p)g^*$). Under $M$, the low type IC can bind without having high types give a concession to low types. The equilibrium separating gift $g^{*M}$ will be larger than $g^*$ in this case, specifically 
\be
g^{*M}= \frac{p(X_{FT}^l-X_{FF}^l)}{1-p}
\ee
High type utility under $M$ is 
\be
U_h^M= pX_{TT}+(1-p)X_{FF}
\ee
because high types exchange concessions of the same value. $U_h^M>U_h^*$. Note that no additional wealth is created by $M$, but less is transferred to low types, and welfare is defined as high-type utility.\footnote{Low types IR constraint still binds and $U_l=X_{FF}$}  $\blacksquare$

The low type leader can bluff in the bilateral setting but not when a mediator is present. As per Theorem 4, inefficiency is due to the high types giving an inefficient gift in order to separate while still being able to prevent low types from using concessions against the high types. But because the benefits of giving a concession differ between the types (low types get one period of cheater's payoff, while high types get future peace), the mediator is able to mandate concessions only between declared high types. This means there is no longer a need for inefficient concessions since concessions will never be used against the giver. Thus, a mediator is able to remove inefficiency in concession giving and bring willing parties to the table to give the gifts necessary to achieve peace.
